quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-05-12 paper-quollr.Rmd
Find the cross reference

Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and (hyper)parameter choices (Figure @fig-nldrlayouts). It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.

@@

```{r fig-nldrlayouts, fig.alt = "NLDR representations with different methods and (hyper)parameter choices.", fig.cap = "Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices. Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical (hyper)parameter choice. Would their downstream analysis decisions vary based on the version they observe? Which representation most accurately represents the structure in high-dimensional space?", out.width = "100%", fig.width= 6, fig.height= 4}

@@

@@

<!-- Assessing which of the 8 NLDR layouts on the limb3k data  (shown in @fig-NLDR-variety) is the better representation using MSE for varying binwidth ($a_1$). Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h). Layout f is universally poor. Layouts a, c, g, h that show large separations between clusters are universally suboptimal. Layout d with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases. The choice of best is between layouts b and e, that have small separations between oddly shaped clusters. Layout e is the best choice.-->

```{r}
#| fig-cap: "Assessing which of the 6 NLDR layouts on the limb3k data  (shown in @fig-nldrlayouts) is the better representation using RMSE for varying binwidth ($a_1$). Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f)."
#| label: fig-limb-mse
#| out-width: 100%

>>>>

Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \@ref(fig:nldr-layouts)). It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.

@@

```{r nldr-layouts, fig.alt = "NLDR representations with different methods and (hyper)parameter choices.", fig.cap = "Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices. Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical (hyper)parameter choice. Would their downstream analysis decisions vary based on the version they observe? Which representation most accurately represents the structure in high-dimensional space?", out.width = "100%", fig.width= 6, fig.height= 4}

@@

\@ref(fig:limb-rmse)

@@

```{r limb-rmse, fig.cap="Assessing which of the 6 NLDR layouts on the limb3k data  (shown in @fig-nldrlayouts) is the better representation using RMSE for varying binwidth ($a_1$). Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f). Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases. Layout f is the best choice.", out.width="100%"}

### 2025-05-12 paper-quollr.tex
Added langevitour of the model
Added tSNE best
Find the cross reference

Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \citet{fig-nldrlayouts}). It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.

@@

\begin{figure}
\includegraphics[width=1\linewidth,alt={NLDR representations with different methods and (hyper)parameter choices.}]{paper-quollr_files/figure-latex/fig-nldrlayouts-1} \caption{Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices. Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical (hyper)parameter choice. Would their downstream analysis decisions vary based on the version they observe? Which representation most accurately represents the structure in high-dimensional space?}\label{fig:fig-nldrlayouts}
\end{figure}

@@

@@

\begin{figure}
\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/fig-limb-mse-1} \end{figure}

>>>>

Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \ref{fig:nldr-layouts}). It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.

@@

\begin{figure}
\includegraphics[width=1\linewidth,alt={NLDR representations with different methods and (hyper)parameter choices.}]{paper-quollr_files/figure-latex/nldr-layouts-1} \caption{Six different NLDR representations of Single-cell RNA sequencing (scRNA-seq) data from mouse limb muscles have been generated using different techniques and different (hyper)parameter choices. Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical (hyper)parameter choice. Would their downstream analysis decisions vary based on the version they observe? Which representation most accurately represents the structure in high-dimensional space?}\label{fig:nldr-layouts}
\end{figure}

@@

#> $nldr_obj
#> $nldr_obj$scaled_nldr
#> # A tibble: 5,000 x 3
#>     emb1  emb2    ID
#>    <dbl> <dbl> <int>
#>  1 0.707 0.839     1
#>  2 0.231 0.401     2
#>  3 0.232 0.215     3
#>  4 0.790 0.564     4
#>  5 0.761 0.551     5
#>  6 0.445 0.721     6
#>  7 0.900 0.137     7
#>  8 0.247 0.392     8
#>  9 0.325 0.542     9
#> 10 0.278 0.231    10
#> $nldr_obj$lim1
#> [1] -14.42166  13.32655
#> $nldr_obj$lim2
#> [1] -12.43687  12.32455

@@

\ref{fig:limb-rmse}

\begin{figure}
\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/limb-rmse-1} \caption{Assessing which of the 6 NLDR layouts on the limb3k data  (shown in @fig-nldrlayouts) is the better representation using RMSE for varying binwidth ($a_1$). Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f). Layout d is perform well at large binwidth (where the binwidth is not enough to capture the data struture) and poorly as bin width decreases. Layout f is the best choice.}\label{fig:limb-rmse}
\end{figure}

\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-28-1}
